---
title: "Assay validation analysis"
output: 
  bookdown::html_document2:
    fig_caption: yes
    theme: spacelab
    toc: yes
    toc_depth: 2
params:
  dfASSAY : NULL
  nBeta : NULL
  nACC_LIMIT : NULL
  name : NULL 
  firstname : NULL
  substance : NULL
  pharmprep : NULL
  concunit : NULL
  signalunit : NULL
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = FALSE)
#require(readr)
require(dplyr) 
#require(rlm)
#require(datalist)
#require(modelr)
require(data.table)
#require(bblme)
require(ggplot2)
require(kableExtra)
require(plotly)
require(viridis)
require(ggpmisc)
require(readxl)
require(papeR)
require(lmtest)
require(pander)
#knitr::write_bib(c(.packages()), file = "packages.bib")


#dfASSAY <- read_excel("data_valid.xlsx")
#dfASSAY <- as.data.table(dfASSAY)
nBETA <- 0.8 
cUNIT <- "(ng/mL)"
nACC_LIMIT <- 10

#source("../R/SCRIPT.R", local = knitr::knit_global())

```

```{=html}
<!-- Reste à mettre les méthodes et calculs
ANOVA sur les profils de linarité
Faire des copier coller
Faciliter l'interprétation -->
```
# Introduction

When writing a paper using a method validated by this software, we recommend that you cite the software beyond and provide the raw data used as supplementary material for your paper.

To cite this software :

> ValidR - Assay

# Methods

## Statistical analysis

All statistical were produced using the method described in Hubert, P. et al. 2007. Harmonization of strategies for the validation of quantitative analytical procedures. A SFSTP proposal - part III. J Pharm Biomed Anal 45: 82-96. And algorithm were checked using the data providied with the article.

The $\beta$ value used for the calculation of the $\beta$ expectation tolerance interval was set to `r paste(nBETA)` and the acceptance limits was fixed to `r paste(nACC_LIMIT,"%")`.

The estimation of the parameters of the calibration curves was obtained by the ordinary least squares method (OLS) with `r paste(R.Version()$version.string)` software [@R-base] and the `r knitr::inline_expr('stats::lm()')` function was used. <!-- This reports was produced using pacakges [@R-data.table;@R-dplyr;@R-ggplot2;@R-ggpmisc;@R-ggpp;@R-kableExtra;@R-plotly;@R-readxl;@R-viridis;@R-viridisLite] -->.

# Analysis of response function (calibration curves)

## Methods and Data

Response functions $signal=f(concentration)$ were analyzed using `stats4::lm()` function in R, using calibration data provided Table \@ref(tab:calibdata) and Figure \@ref(fig:calibdraw)).

```{r, warning = FALSE, echo=FALSE}
calib_curves <- data.frame("method"=character(),
                            "series"=character(),
                            "intercept"=numeric(),
                            "slope"=numeric(),
                            "aic"=numeric(),
                            "r2"=numeric())

diag_calib <- data.frame(resid=numeric())

compute_calibration_curves(dfASSAY)
```

Analysis were performed using :

-   Linear function (Linear)
-   Linear trough 0 function (Linear 0)
-   Linear trough 0 and the highest concentration levels function (Linear 0 - Max)
-   Linear function with weights (1/Y and 1/X) (Linear weighed 1/X and Linear weighed 1/Y)

```{r calibdraw, echo=FALSE, fig.cap="Calibration data", fig.fullwidth=TRUE}

ggplotly(print_plot_standards(dfASSAY, "CAL"))

```

## Response functions obtained

### Regression analysis

The interactive table \@ref(tab:resultscalib) shows the values obtained with regressions :

```{r resultscalib, echo=FALSE}

calib_curves %>% arrange(method) %>%
                 kableExtra::kable(col.names = c('Methods', 'Serie', 'Intercept', 'Slope', 'AIC', "R²"), 
                                   caption = "Results of linear regressions performed") %>%
                             kable_styling("hover") 

```

### Residues for each calibration curves at each levels

The residues obtained are shown in the interactive figure \@ref(fig:calibresidues).

You can click or double-click on a legend to isolate or display a curve. You can zoom in on a part of the figure.

```{r calibresidues, echo=FALSE, fig.width=10,fig.height=4,fig.fullwidth=TRUE, fig.cap="Relative bias calculated from regression"}

ggplotly(plot_residues(dfASSAY))

```

# Validation

```{r, echo=FALSE}

compute_validation_data(dfASSAY)

```

## Using a linear calibration curve

### Trueness and precision obtained

Trueness and precision are depicted on table \@ref(tab:iplinabs) and \@ref(tab:iplinrel)

```{r iplinabs, echo=FALSE, warning=FALSE, fig.width=10,fig.height=4,fig.fullwidth=TRUE, fig.cap="Trueness and precision obtained"}

summary_validation_abs(as.data.frame(listVALID[["LIN"]]), names(listVALID["LIN"]))

```

```{r iplinrel, echo=FALSE, warning=FALSE, fig.width=10,fig.height=4,fig.fullwidth=TRUE, fig.cap="Trueness and precision obtained"}

summary_validation_rel(as.data.frame(listVALID[["LIN"]]), names(listVALID["LIN"]))

```

### Accuracy profile

Accuracy profile are shown Figure \@ref(fig:aplin). The 𝛽-tolerance interval (blue lines) should be entirely within the acceptance limits (red dashed lines).

```{r aplin, echo=FALSE, fig.width=10, fig.height=4, fig.fullwidth=TRUE, fig.cap="Accuracy profiles  (red dashed line: acceptance limits, blue lines: 𝛽-tolerance intervals)."}

  ggplotly(plot_validation(as.data.frame(listVALID[["LIN"]]), names(listVALID["LIN"])))


```

### Linearity profile

Linearity profile are shown Figure \@ref(fig:lplin). You should check that the black line (linear model) should be superimposed on the red dashed identity line.

```{r lplin, echo=FALSE, fig.width=10, fig.height=4, fig.fullwidth=TRUE, fig.cap="Linearity profiles (red dashed line: identity line, black lines: linear regression lines, blue lines: 𝛽-tolerance intervals)."}

  ggplotly(plot_linearity(as.data.frame(listVALID[["LIN"]]), names(listVALID["LIN"])))

```

### Linear regression

`r if(summary(linear_model_stat(listVALID[["LIN"]]))$coefficients[1,4]<0.05){ "The p-value for the intercept is < 0.05 which is in favor of a significant value (FAIL)"}else{ "The p-value for the intercept is > 0.05 which is in favor of a non-significant value (PASS)"}`

`r if(summary(linear_model_stat(listVALID[["LIN"]]))$coefficients[2,4]<0.05){ "The p-value for the slope is < 0.05 which is in favor of a significant value (PASS)"}else{ "The p-value for the slope is > 0.05 which is in favor of a non-significant value (FAIL)"}`

```{r}

kbl(prettify(summary(linear_model_stat(listVALID[["LIN"]]))), 
    caption = 'Results of the linear regression.') %>%
  kable_styling(full_width = F, position = "left") 
```

### Studentized Breusch-Pagan test for heteroskedasticity

`r if(bptest(linear_model_stat(listVALID[["LIN"]]))$p.value<0.05){"The p-value < 0.05 which is in favor of heterosedasticity and standard deviations of a calculated concentrations as related to introduced concentrations, are non-constant (FAIL)"}else{"The p-value > O.05 which is in favor of homosedasticity and standard deviations of a calculated concentrations as related to introduced concentrations, are constant (PASS)"}`

<!-- is in favor of heterosedasticity (the standard deviations of a calculated concentrations as related to introduced concentrations, are non-constant)"} else {"A p-value > 0.05 is in favor of homosedasticity (the standard deviations of a calculated concentrations as related to introduced concentrations, are constant")}` -->

```{r }

pander(bptest(linear_model_stat(listVALID[["LIN"]])), caption="Result of the studentized Breusch-Pagan test")

```

## Using a linear forced trough 0 calibration curve

### Trueness and precision obtained

Trueness and precision are depicted on table \@ref(tab:iplinabs) and \@ref(tab:iplinrel)

```{r iplin0abs, echo=FALSE, warning=FALSE, fig.width=10,fig.height=4,fig.fullwidth=TRUE, fig.cap="Trueness and precision obtained"}

summary_validation_abs(as.data.frame(listVALID[["LIN_0"]]), names(listVALID["LIN_0"]))

```

```{r iplin0rel, echo=FALSE, warning=FALSE, fig.width=10,fig.height=4,fig.fullwidth=TRUE, fig.cap="Trueness and precision obtained"}

summary_validation_rel(as.data.frame(listVALID[["LIN_0"]]), names(listVALID["LIN_0"]))

```

### Accuracy profile

Accuracy profile are shown Figure \@ref(fig:aplin). The 𝛽-tolerance interval (blue lines) should be entirely within the acceptance limits (red dashed lines).

```{r aplin0, echo=FALSE, fig.width=10, fig.height=4, fig.fullwidth=TRUE, fig.cap="Accuracy profiles  (red dashed line: acceptance limits, blue lines: 𝛽-tolerance intervals)."}

  ggplotly(plot_validation(as.data.frame(listVALID[["LIN_0"]]), names(listVALID["LIN_0"])))


```

### Linearity profile

Linearity profile are shown Figure \@ref(fig:lplin). You should check that the black line (linear model) should be superimposed on the red dashed identity line.

```{r lplin0, echo=FALSE, fig.width=10, fig.height=4, fig.fullwidth=TRUE, fig.cap="Linearity profiles (red dashed line: identity line, black lines: linear regression lines, blue lines: 𝛽-tolerance intervals)."}

  ggplotly(plot_linearity(as.data.frame(listVALID[["LIN_0"]]), names(listVALID["LIN_0"])))

```

### Linear regression

`r if(summary(linear_model_stat(listVALID[["LIN_0"]]))$coefficients[1,4]<0.05){ "The p-value for the intercept is < 0.05 which is in favor of a significant value (FAIL)"}else{ "The p-value for the intercept is > 0.05 which is in favor of a non-significant value (PASS)"}`

`r if(summary(linear_model_stat(listVALID[["LIN_0"]]))$coefficients[2,4]<0.05){ "The p-value for the slope is < 0.05 which is in favor of a significant value (PASS)"}else{ "The p-value for the slope is > 0.05 which is in favor of a non-significant value (FAIL)"}`

```{r}

kbl(prettify(summary(linear_model_stat(listVALID[["LIN_0"]]))), 
    caption = 'Results of the linear regression.') %>%
  kable_styling(full_width = F, position = "left") 
```

### Studentized Breusch-Pagan test for heteroskedasticity

`r if(bptest(linear_model_stat(listVALID[["LIN_0"]]))$p.value<0.05){"The p-value < 0.05 which is in favor of heterosedasticity and standard deviations of a calculated concentrations as related to introduced concentrations, are non-constant (FAIL)"}else{"The p-value > O.05 which is in favor of homosedasticity and standard deviations of a calculated concentrations as related to introduced concentrations, are constant (PASS)"}`

```{r }

pander(bptest(linear_model_stat(listVALID[["LIN_0"]])), caption="Result of the studentized Breusch-Pagan test")

```

## Using a linear calibration curve with the highest level only and forced trough 0

### Trueness and precision obtained

Trueness and precision are depicted on table \@ref(tab:iplinabs) and \@ref(tab:iplinrel)

```{r iplinmaxabs, echo=FALSE, warning=FALSE, fig.width=10,fig.height=4,fig.fullwidth=TRUE, fig.cap="Trueness and precision obtained"}

summary_validation_abs(as.data.frame(listVALID[["LIN_0MAX"]]), names(listVALID["LIN_0MAX"]))

```

```{r iplinmaxrel, echo=FALSE, warning=FALSE, fig.width=10,fig.height=4,fig.fullwidth=TRUE, fig.cap="Trueness and precision obtained"}

summary_validation_rel(as.data.frame(listVALID[["LIN_0MAX"]]), names(listVALID["LIN_0MAX"]))

```

### Accuracy profile

Accuracy profile are shown Figure \@ref(fig:aplin). The 𝛽-tolerance interval (blue lines) should be entirely within the acceptance limits (red dashed lines).

```{r aplinmax, echo=FALSE, fig.width=10, fig.height=4, fig.fullwidth=TRUE, fig.cap="Accuracy profiles  (red dashed line: acceptance limits, blue lines: 𝛽-tolerance intervals)."}

  ggplotly(plot_validation(as.data.frame(listVALID[["LIN_0MAX"]]), names(listVALID["LIN_0MAX"])))


```

### Linearity profile

Linearity profile are shown Figure \@ref(fig:lplin). You should check that the black line (linear model) should be superimposed on the red dashed identity line.

```{r lplinmax, echo=FALSE, fig.width=10, fig.height=4, fig.fullwidth=TRUE, fig.cap="Linearity profiles (red dashed line: identity line, black lines: linear regression lines, blue lines: 𝛽-tolerance intervals)."}

  ggplotly(plot_linearity(as.data.frame(listVALID[["LIN_0MAX"]]), names(listVALID["LIN_0MAX"])))

```

### Linear regression

`r if(summary(linear_model_stat(listVALID[["LIN_0MAX"]]))$coefficients[1,4]<0.05){ "The p-value for the intercept is < 0.05 which is in favor of a significant value (FAIL)"}else{ "The p-value for the intercept is > 0.05 which is in favor of a non-significant value (PASS)"}`

`r if(summary(linear_model_stat(listVALID[["LIN_0MAX"]]))$coefficients[2,4]<0.05){ "The p-value for the slope is < 0.05 which is in favor of a significant value (PASS)"}else{ "The p-value for the slope is > 0.05 which is in favor of a non-significant value (FAIL)"}`

```{r}

kbl(prettify(summary(linear_model_stat(listVALID[["LIN_0MAX"]]))), 
    caption = 'Results of the linear regression.') %>%
  kable_styling(full_width = F, position = "left") 
```

### Studentized Breusch-Pagan test for heteroskedasticity

`r if(bptest(linear_model_stat(listVALID[["LIN_0MAX"]]))$p.value<0.05){"The p-value < 0.05 which is in favor of heterosedasticity and standard deviations of a calculated concentrations as related to introduced concentrations, are non-constant (FAIL)"}else{"The p-value > O.05 which is in favor of homosedasticity and standard deviations of a calculated concentrations as related to introduced concentrations, are constant (PASS)"}`

<!-- is in favor of heterosedasticity (the standard deviations of a calculated concentrations as related to introduced concentrations, are non-constant)"} else {"A p-value > 0.05 is in favor of homosedasticity (the standard deviations of a calculated concentrations as related to introduced concentrations, are constant")}` -->

```{r }

pander(bptest(linear_model_stat(listVALID[["LIN_0MAX"]])), caption="Result of the studentized Breusch-Pagan test")

```

## Using a weighed (1/Y) linear calibration curve

### Trueness and precision obtained

Trueness and precision are depicted on table \@ref(tab:iplinabs) and \@ref(tab:iplinrel)

```{r iplin1yabs, echo=FALSE, warning=FALSE, fig.width=10,fig.height=4,fig.fullwidth=TRUE, fig.cap="Trueness and precision obtained"}

summary_validation_abs(as.data.frame(listVALID[["LIN_1Y"]]), names(listVALID["LIN_1Y"]))

```

```{r iplin1yrel, echo=FALSE, warning=FALSE, fig.width=10,fig.height=4,fig.fullwidth=TRUE, fig.cap="Trueness and precision obtained"}

summary_validation_rel(as.data.frame(listVALID[["LIN_1Y"]]), names(listVALID["LIN_1Y"]))

```

### Accuracy profile

Accuracy profile are shown Figure \@ref(fig:aplin). The 𝛽-tolerance interval (blue lines) should be entirely within the acceptance limits (red dashed lines).

```{r aplin1y, echo=FALSE, fig.width=10, fig.height=4, fig.fullwidth=TRUE, fig.cap="Accuracy profiles  (red dashed line: acceptance limits, blue lines: 𝛽-tolerance intervals)."}

  ggplotly(plot_validation(as.data.frame(listVALID[["LIN_1Y"]]), names(listVALID["LIN_1Y"])))


```

### Linearity profile

Linearity profile are shown Figure \@ref(fig:lplin). You should check that the black line (linear model) should be superimposed on the red dashed identity line.

```{r lplin1y, echo=FALSE, fig.width=10, fig.height=4, fig.fullwidth=TRUE, fig.cap="Linearity profiles (red dashed line: identity line, black lines: linear regression lines, blue lines: 𝛽-tolerance intervals)."}

  ggplotly(plot_linearity(as.data.frame(listVALID[["LIN_1Y"]]), names(listVALID["LIN_1Y"])))

```

### Linear regression

`r if(summary(linear_model_stat(listVALID[["LIN_1Y"]]))$coefficients[1,4]<0.05){ "The p-value for the intercept is < 0.05 which is in favor of a significant value (FAIL)"}else{ "The p-value for the intercept is > 0.05 which is in favor of a non-significant value (PASS)"}`

`r if(summary(linear_model_stat(listVALID[["LIN_1Y"]]))$coefficients[2,4]<0.05){ "The p-value for the slope is < 0.05 which is in favor of a significant value (PASS)"}else{ "The p-value for the slope is > 0.05 which is in favor of a non-significant value (FAIL)"}`

```{r}

kbl(prettify(summary(linear_model_stat(listVALID[["LIN_1Y"]]))), 
    caption = 'Results of the linear regression.') %>%
  kable_styling(full_width = F, position = "left") 
```

### Studentized Breusch-Pagan test for heteroskedasticity

`r if(bptest(linear_model_stat(listVALID[["LIN_1Y"]]))$p.value<0.05){"The p-value < 0.05 which is in favor of heterosedasticity and standard deviations of a calculated concentrations as related to introduced concentrations, are non-constant (FAIL)"}else{"The p-value > O.05 which is in favor of homosedasticity and standard deviations of a calculated concentrations as related to introduced concentrations, are constant (PASS)"}`

```{r }

pander(bptest(linear_model_stat(listVALID[["LIN_1Y"]])), caption="Result of the studentized Breusch-Pagan test")

```

## Using a weighed (1/X) linear calibration curve

### Trueness and precision obtained

Trueness and precision are depicted on table \@ref(tab:iplinabs) and \@ref(tab:iplinrel)

```{r iplin1xabs, echo=FALSE, warning=FALSE, fig.width=10,fig.height=4,fig.fullwidth=TRUE, fig.cap="Trueness and precision obtained"}

summary_validation_abs(as.data.frame(listVALID[["LIN_1X"]]), names(listVALID["LIN_1X"]))

```

```{r iplin1xrel, echo=FALSE, warning=FALSE, fig.width=10,fig.height=4,fig.fullwidth=TRUE, fig.cap="Trueness and precision obtained"}

summary_validation_rel(as.data.frame(listVALID[["LIN_1X"]]), names(listVALID["LIN_1X"]))

```

### Accuracy profile

Accuracy profile are shown Figure \@ref(fig:aplin). The 𝛽-tolerance interval (blue lines) should be entirely within the acceptance limits (red dashed lines).

```{r aplin1x, echo=FALSE, fig.width=10, fig.height=4, fig.fullwidth=TRUE, fig.cap="Accuracy profiles  (red dashed line: acceptance limits, blue lines: 𝛽-tolerance intervals)."}

  ggplotly(plot_validation(as.data.frame(listVALID[["LIN_1X"]]), names(listVALID["LIN_1X"])))


```

### Linearity profile

Linearity profile are shown Figure \@ref(fig:lplin). You should check that the black line (linear model) should be superimposed on the red dashed identity line.

```{r lplin1x, echo=FALSE, fig.width=10, fig.height=4, fig.fullwidth=TRUE, fig.cap="Linearity profiles (red dashed line: identity line, black lines: linear regression lines, blue lines: 𝛽-tolerance intervals)."}

  ggplotly(plot_linearity(as.data.frame(listVALID[["LIN_1X"]]), names(listVALID["LIN_1X"])))

```

### Linear regression

`r if(summary(linear_model_stat(listVALID[["LIN_1X"]]))$coefficients[1,4]<0.05){ "The p-value for the intercept is < 0.05 which is in favor of a significant value (FAIL)"}else{ "The p-value for the intercept is > 0.05 which is in favor of a non-significant value (PASS)"}`

`r if(summary(linear_model_stat(listVALID[["LIN_1X"]]))$coefficients[2,4]<0.05){ "The p-value for the slope is < 0.05 which is in favor of a significant value (PASS)"}else{ "The p-value for the slope is > 0.05 which is in favor of a non-significant value (FAIL)"}`

```{r}

kbl(prettify(summary(linear_model_stat(listVALID[["LIN_1X"]]))), 
    caption = 'Results of the linear regression.') %>%
  kable_styling(full_width = F, position = "left") 
```

### Studentized Breusch-Pagan test for heteroskedasticity

`r if(bptest(linear_model_stat(listVALID[["LIN_1X"]]))$p.value<0.05){"The p-value < 0.05 which is in favor of heterosedasticity and standard deviations of a calculated concentrations as related to introduced concentrations, are non-constant (FAIL)"}else{"The p-value > O.05 which is in favor of homosedasticity and standard deviations of a calculated concentrations as related to introduced concentrations, are constant (PASS)"}`

```{r }

pander(bptest(linear_model_stat(listVALID[["LIN_1X"]])), caption="Result of the studentized Breusch-Pagan test")

```

# Raw data

```{r calibdata, echo=FALSE}

print_table_rawdata(dfASSAY, "CAL")

```

```{r validdata, echo=FALSE}

print_table_rawdata(dfASSAY, "VAL")

```

# Packages

Aphalo, Pedro J. 2022a. Ggpmisc: Miscellaneous Extensions to Ggplot2. https://CRAN.R-project.org/package=ggpmisc.

Aphalo, Pedro J. 2022b. Ggpp: Grammar Extensions to Ggplot2. https://CRAN.R-project.org/package=ggpp.

Dahl, David B., David Scott, Charles Roosen, Arni Magnusson, and Jonathan Swinton. 2019. Xtable: Export Tables to LaTeX or HTML. http://xtable.r-forge.r-project.org/.

Daróczi, Gergely, and Roman Tsegelskyi. 2022. Pander: An r Pandoc Writer. https://rapporter.github.io/pander/.

Dowle, Matt, and Arun Srinivasan. 2022. Data.table: Extension of ‘Data.frame‘. https://CRAN.R-project.org/package=data.table.

Fox, John, and Sanford Weisberg. 2019. An R Companion to Applied Regression. Third. Thousand Oaks CA: Sage. https://socialsciences.mcmaster.ca/jfox/Books/Companion/.

Fox, John, Sanford Weisberg, and Brad Price. 2022a. Car: Companion to Applied Regression. https://CRAN.R-project.org/package=car.

Fox, John, Sanford Weisberg, and Brad Price. 2022b. carData: Companion to Applied Regression Data Sets. https://CRAN.R-project.org/package=carData.

Garnier, Simon. 2021. Viridis: Colorblind-Friendly Color Maps for r. https://CRAN.R-project.org/package=viridis.

Garnier, Simon.. 2022. viridisLite: Colorblind-Friendly Color Maps (Lite Version). https://CRAN.R-project.org/package=viridisLite.

Hofner, Benjamin. 2021. papeR: A Toolbox for Writing Pretty Papers and Reports. https://CRAN.R-project.org/package=papeR.

Hofner, Benjamin, and with contributions by many others. 2021. papeR: A Toolbox for Writing Pretty Papers and Reports. https://github.com/hofnerb/papeR.

Hothorn, Torsten, Achim Zeileis, Richard W. Farebrother, and Clint Cummins. 2022. Lmtest: Testing Linear Regression Models. https://CRAN.R-project.org/package=lmtest.

R Core Team. 2022. R: A Language and Environment for Statistical Computing. Vienna, Austria: R Foundation for Statistical Computing. https://www.R-project.org/.

Sievert, Carson. 2020. Interactive Web-Based Data Visualization with r, Plotly, and Shiny. Chapman; Hall/CRC. https://plotly-r.com.

Sievert, Carson, Chris Parmer, Toby Hocking, Scott Chamberlain, Karthik Ram, Marianne Corvellec, and Pedro Despouy. 2022. Plotly: Create Interactive Web Graphics via Plotly.js. https://CRAN.R-project.org/package=plotly.

Wickham, Hadley. 2016. Ggplot2: Elegant Graphics for Data Analysis. Springer-Verlag New York. https://ggplot2.tidyverse.org.

Wickham, Hadley, and Jennifer Bryan. 2022. Readxl: Read Excel Files. https://CRAN.R-project.org/package=readxl.

Wickham, Hadley, Winston Chang, Lionel Henry, Thomas Lin Pedersen, Kohske Takahashi, Claus Wilke, Kara Woo, Hiroaki Yutani, and Dewey Dunnington. 2022. Ggplot2: Create Elegant Data Visualisations Using the Grammar of Graphics. https://CRAN.R-project.org/package=ggplot2.

Wickham, Hadley, Romain François, Lionel Henry, and Kirill Müller. 2022. Dplyr: A Grammar of Data Manipulation. https://CRAN.R-project.org/package=dplyr.

Zeileis, Achim, and Gabor Grothendieck. 2005. “Zoo: S3 Infrastructure for Regular and Irregular Time Series.” Journal of Statistical Software 14 (6): 1–27. https://doi.org/10.18637/jss.v014.i06.

Zeileis, Achim, Gabor Grothendieck, and Jeffrey A. Ryan. 2022. Zoo: S3 Infrastructure for Regular and Irregular Time Series (z’s Ordered Observations). https://zoo.R-Forge.R-project.org/.

Zeileis, Achim, and Torsten Hothorn. 2002. “Diagnostic Checking in Regression Relationships.” R News 2 (3): 7–10. https://CRAN.R-project.org/doc/Rnews/.

Zhu, Hao. 2021. kableExtra: Construct Complex Table with Kable and Pipe Syntax. https://CRAN.R-project.org/package=kableExtra.
